---
title: "Project Assignment 4: Clustering"
author: "XINYI LIU, Zhijiao Chen, Po Chiao Chen, Yi Chen"
date: "3/02/2020"
output:
  html_document: default
  word_document: default
---
```{r, warning=FALSE}
library(dplyr)
library(tidyverse)
library(forecast)
library(leaps)
library(forecast)
library(caret)
library(car)
library(data.table)
```


```{r}
weather_train <- read.csv("weather_train.csv")
building_metadata <- read.csv("building_metadata.csv")
train <- fread("train.csv")

weather_train1 <- weather_train %>% filter(weather_train$site_id == 4)
weather_train1$timestamp <- as.Date(weather_train1$timestamp)
weather <- weather_train1 %>% group_by(timestamp = weather_train1$timestamp) %>% summarise(avg_temp = avg_temp <- mean(air_temperature,na.rm = TRUE),avg_cloud = avg_cloud <- mean(cloud_coverage,na.rm = TRUE),avg_dewtemp = avg_dewtemp <- mean(dew_temperature,na.rm = TRUE),avg_sea = avg_sea <- mean(sea_level_pressure,na.rm = TRUE),avg_wind = avg_wind <- mean(wind_speed,na.rm = TRUE))


```

```{r}
building <- dplyr::inner_join(building_metadata,train,by="building_id")
building <- building %>% filter(building$site_id == 4)
building$timestamp <- as.Date(building$timestamp)
building_group <- building %>% group_by(building_id = building$building_id, timestamp = building$timestamp) %>%
  summarise(avg_meter_reading = avg_meter_reading <- mean(meter_reading,na.rm=TRUE), meter = meter <- mean(meter)) 

#building_filter_site4 <- building_metadata %>% filter(building_metadata$site_id == 4)
building_final <- dplyr::inner_join(building_group, building_metadata,by="building_id")
final_df <- dplyr::inner_join(weather,building_final,by="timestamp")

#unique(final_df$meter)
#hist(train$meter)
#sum(is.na(final_df))
``` 

```{r}
final_df['age']=2016-final_df$year_built
final_df <- final_df[,-13]
```

# Omiting NA
```{r}
final_df <- na.omit(final_df)
```

# Normalize data 
```{r, fig.height= 10}
par(mfrow = c(2, 3))
final_df_normalize <- final_df %>% mutate(
                    avg_temp = scale(final_df$avg_temp), 
                    avg_cloud = scale(final_df$avg_cloud), 
                    avg_dewtemp = scale(final_df$avg_dewtemp),
                    avg_sea = scale(final_df$avg_sea),
                    avg_wind = scale(final_df$avg_wind),
                    avg_meter_reading = scale(final_df$avg_meter_reading),
                    square_feet = scale(final_df$square_feet),
                    floor_count = scale(final_df$floor_count),
                    age = scale(final_df$age)
                    )
hist(final_df_normalize$avg_temp, density = 20, breaks = 20, prob = TRUE, main = "Avg Temp Histogram", cex.main =3, cex.axis = 2.5, cex.lab = 2)
curve(dnorm, add = TRUE, col = "black", lwd = 5)

hist(final_df_normalize$avg_cloud, density = 20, breaks = 20, prob = TRUE, main = "Avg Cloud Histogram", cex.main =3, cex.axis = 2.5, cex.lab = 2)
curve(dnorm, add = TRUE, col = "black", lwd = 5)

hist(final_df_normalize$avg_dewtemp, density = 20, breaks = 20, prob = TRUE, main = "Avg Dewtemp Histogram", cex.main =3, cex.axis = 2.5, cex.lab = 2)
curve(dnorm, add = TRUE, col = "black", lwd = 5)

hist(final_df_normalize$avg_sea, density = 20, breaks = 20, prob = TRUE, main = "Avg Sea Histogram", cex.main =3, cex.axis = 2.5, cex.lab = 2)
curve(dnorm, add = TRUE, col = "black", lwd = 5)

hist(final_df_normalize$avg_wind, density = 20, breaks = 20, prob = TRUE, main = "Avg Wind Histogram", cex.main =3, cex.axis = 2.5, cex.lab = 2)
curve(dnorm, add = TRUE, col = "black", lwd = 5)

hist(final_df_normalize$avg_meter_reading, density = 20, breaks = 20, prob = TRUE, main = "Avg Meter Reading Histogram", cex.main =3, cex.axis = 2.5, cex.lab = 2)
curve(dnorm, add = TRUE, col = "black", lwd = 5)

hist(final_df_normalize$square_feet, density = 20, breaks = 20, prob = TRUE, main = "Avg Square Feet Histogram", cex.main =3, cex.axis = 2.5, cex.lab = 2)
curve(dnorm, add = TRUE, col = "black", lwd = 5)

hist(final_df_normalize$floor_count, density = 20, breaks = 20, prob = TRUE, main = "Avg Floor Count Histogram", cex.main =3, cex.axis = 2.5, cex.lab = 2)
curve(dnorm, add = TRUE, col = "black", lwd = 5)

hist(final_df_normalize$age, density = 20, breaks = 20, prob = TRUE, main = "Avg Building Age Histogram", cex.main =3, cex.axis = 2.5, cex.lab = 2)
curve(dnorm, add = TRUE, col = "black", lwd = 5)

hist(final_df$meter, prob = TRUE, main = "Meter Type Histogram", cex.main =3, cex.axis = 2.5, cex.lab = 2)

hist(as.numeric(final_df$primary_use), density = 20, breaks = 20, prob = TRUE,main = "Avg Primary Use Histogram", cex.main =3, cex.axis = 2.5, cex.lab = 2)
curve(dnorm, add = TRUE, col = "black", lwd = 5)
```

### 1. Show descriptive statistics for relevant and important variables, particularly for your target variable and all your model’s candidate predictor variables.
box-and-whisker plots for relevant and important variables
```{r,fig.width=7,fig.height=7}

par(mfrow=c(1,6))
boxplot(final_df$avg_temp,main="Temperature")
boxplot(final_df$avg_cloud,main="Cloud Coverage")
boxplot(final_df$avg_dewtemp,main="Dew Temperature")
boxplot(final_df$avg_sea,main="Sea Level")
boxplot(final_df$avg_wind,main="Wind Speed")
boxplot(final_df$avg_meter_reading,main="Meter Reading")
boxplot(final_df$meter,main="Meter Type")
boxplot(as.numeric(final_df$primary_use),main="Building Type")
boxplot(final_df$square_feet,main="Building Square feet")
boxplot(final_df$floor_count,main="Building Floor")
boxplot(final_df$age,main="Busilding Age")

```

The minimum, maximum, and average (mean, median, mode) and standard deviation / variance of important variables. 
```{r}
library(kableExtra)
Minimum <-c(min(final_df$avg_temp),min(final_df$avg_cloud),min(final_df$avg_dewtemp),
            min(final_df$avg_sea),min(final_df$avg_wind),min(final_df$avg_meter_reading),
            min(final_df$meter),min(as.numeric(final_df$primary_use)), min(final_df$square_feet), 
            min(final_df$floor_count), min(final_df$age))

Maximum <- c(max(final_df$avg_temp),max(final_df$avg_cloud),max(final_df$avg_dewtemp),
            max(final_df$avg_sea),max(final_df$avg_wind),max(final_df$avg_meter_reading),
            max(final_df$meter),max(as.numeric(final_df$primary_use)), max(final_df$square_feet), 
            max(final_df$floor_count), max(final_df$age))


Mean <-c(mean(final_df$avg_temp),mean(final_df$avg_cloud),mean(final_df$avg_dewtemp),
            mean(final_df$avg_sea),mean(final_df$avg_wind),mean(final_df$avg_meter_reading),
            mean(final_df$meter),mean(as.numeric(final_df$primary_use)), mean(final_df$square_feet), 
            mean(final_df$floor_count), mean(final_df$age))

Median <- c(median(final_df$avg_temp),median(final_df$avg_cloud),median(final_df$avg_dewtemp),
            median(final_df$avg_sea),median(final_df$avg_wind),median(final_df$avg_meter_reading),
            median(final_df$meter),median(as.numeric(final_df$primary_use)), median(final_df$square_feet), 
            median(final_df$floor_count), median(final_df$age))

Sd <- c(sd(final_df$avg_temp),sd(final_df$avg_cloud),sd(final_df$avg_dewtemp),
            sd(final_df$avg_sea),sd(final_df$avg_wind),sd(final_df$avg_meter_reading),
            sd(final_df$meter),sd(as.numeric(final_df$primary_use)), sd(final_df$square_feet), 
            sd(final_df$floor_count), sd(final_df$age)) 
# bring all together
#names(table) <- c("Minimum","Maximum","Average", "Median","Standard Deviation")
all = data.frame(Minimum,Maximum,Mean,Median,Sd)
row.names(all)<-c("Temperature","Cloud Coverage","Dew Temperature","Sea Level","Wind Speed", "Meter Reading","Meter Type", "Building Type","Building Square Feet", "Building Floor","Building Age")
all %>% 
  knitr::kable(digits=3,align='c')%>% 
  kable_styling()

```

### 2. Create a scatterplot among the variables to find potentially linear or curvilinear relationships. That should help you identify both a target variable and candidate predictor variables.  

```{r,fig.width=10,fig.height=10}
library(corrplot)
pairs(final_df[,-c(1,7,9,10,11)])
```

# multiple regression
```{r}
sum(is.na(final_df$age))
reg <- lm(avg_meter_reading ~. - building_id -meter - site_id, data = final_df_normalize)

summary(reg)
vif(reg)
```

## hierarchical clustering
```{r}
final_df_normalize1 <- final_df_normalize
final_df_normalize1 <- final_df_normalize1[,-c(1,7,9,10,11)]
```

```{r}
# compute normalized distance based on all variables
d.norm1 <- dist(final_df_normalize1, method = "euclidean")  
```

```{r}
# in hclust() set LINKAGE parameter to "ward.D", "single", "complete", "average", "median", or "centroid"
hc1 <- hclust(d.norm1, method = "single")
plot(hc1, hang = -1, ann = FALSE) 
# Note: hang -1 means do not hang labels off the leaves; make them level; ann is for plot annotation

hc2 <- hclust(d.norm1, method = "average")
plot(hc2, hang = -1, ann = FALSE)

hc3 <- hclust(d.norm1, method = "median")
plot(hc3, hang = -1, ann = FALSE)

hc4 <- hclust(d.norm1, method = "complete")
plot(hc4, hang = -1, ann = FALSE)

hc5 <- hclust(d.norm1, method = "centroid")
plot(hc5, hang = -1, ann = FALSE)

hc6 <- hclust(d.norm1, method = "ward.D")
plot(hc6, hang = -1, ann = FALSE)
```

```{r}
d.norm2 <- dist(final_df_normalize1, method = "maximum")  
d.norm3 <- dist(final_df_normalize1, method = "manhattan")  
d.norm4 <- dist(final_df_normalize1, method = "canberra") 
d.norm5 <- dist(final_df_normalize1, method = "binary") 
d.norm6 <- dist(final_df_normalize1, method = "minkowski") 
```

```{r}
hc7 <- hclust(d.norm2, method = "ward.D")
plot(hc7, hang = -1, ann = FALSE)
hc8 <- hclust(d.norm3, method = "ward.D")
plot(hc8, hang = -1, ann = FALSE)
hc9 <- hclust(d.norm4, method = "ward.D")
plot(hc9, hang = -1, ann = FALSE)
hc10 <- hclust(d.norm5, method = "ward.D")
plot(hc10, hang = -1, ann = FALSE)
hc11 <- hclust(d.norm6, method = "ward.D")
plot(hc11, hang = -1, ann = FALSE)
```

```{r}
memb <- cutree(hc9, k = 3) 
cat(memb)
hist(memb)
```


```{r}
heatmap(as.matrix(final_df_normalize1), Colv = NA, hclustfun = hclust, 
        col=rev(paste("grey",1:99,sep="")))
```

```{r}
final_df_normalize1['HierCluster'] <- memb
mean_label <- final_df_normalize1 %>% group_by(HierCluster) %>% summarise_each(funs(mean))
mean_label['statistics'] <- 'avg'
max_label <- final_df_normalize1 %>% group_by(HierCluster) %>% summarise_each(funs(max))
max_label['statistics'] <- 'max'
min_label <- final_df_normalize1 %>% group_by(HierCluster) %>% summarise_each(funs(min))
min_label['statistics'] <- 'min'
final_label <- rbind(mean_label,max_label,min_label)
```

```{r}
memb1 <- memb
memb1[memb1 == 2] <- "Top"
memb1[memb1 == 1] <- "medium"
memb1[memb1 == 3] <- "small"
row.names(final_df_normalize1) <- paste(memb1, ": ", row.names(final_df_normalize1), sep = "")
row.names(final_df_normalize1)
```

## K-Means
```{r}
set.seed(1)
final_df_normalize2 <- final_df_normalize[,-c(1,7,9,10,11)]
final_df_normalize2

#Elbow Method for finding the optimal number of clusters
set.seed(123)
# Compute and plot wss for k = 2 to k = 15.
k.max <- 20
data <- final_df_normalize2
wss <- sapply(1:k.max, 
              function(k){kmeans(data, k )$tot.withinss})
wss
plot(1:k.max, wss,
     type="b", pch = 19, frame = FALSE, 
     xlab="Number of clusters K",
     ylab="Total within-clusters sum of squares")
```


```{r}
set.seed(123)
km_final <- kmeans(final_df_normalize2, 3)

# plot an empty scatter plot 
plot(c(0), xaxt = 'n',xlab="", ylab = "", type = "l", ylim = c(min(km_final$centers), max(km_final$centers)), xlim = c(0, 9))
# label x-axes 
axis(1, at = c(1:9), labels = names(final_df_normalize2),par(las="2"))
# plot centroids 
for (i in c(1:3)) 
  lines(km_final$centers[i,], lty = i, lwd = 2, col = ifelse(i %in% c(1,3), "black", "dark grey"))
# name clusters 
text(x = 0.5, y = km_final$centers[, 1], labels = paste("Cluster", c(1:3)))

```

```{r}
final_df_normalize3 <- final_df_normalize2
final_df_normalize3$kmeans <- km_final$cluster
```

```{r}
final <- final_df_normalize3 %>% group_by(kmeans) %>%
  summarise(temp = mean(avg_temp), cloud=mean(avg_cloud),dewtemp=mean(avg_dewtemp),sea=mean(avg_sea),wind=mean(avg_wind),meter=mean(avg_meter_reading),squarefeet=mean(square_feet),age=mean(age))

final %>% 
  knitr::kable(digits=3,align='c')%>% 
  kable_styling()
```
```{r}
a<- km_final$cluster
a[a==1] <- "Large and Warm"
a[a==2] <- "Small and Warm"
a[a==3] <- "Small and Cold"
final_df_normalize3$kmeans <- a
```



```{r}
final_df_normalize3['HierCluster'] <- memb
```



## multiple regression
```{r}
# partition data
set.seed(1)  # set seed for reproducing the partition
final_df_normalize3$primary_use <- final_df_normalize$primary_use
train.index <- sample(c(1:27340), 27340*0.6)  

#Create and set aside the remaining 40% of the data, to be used after omitting unhelpful data points and unnecessary variables.
train.df <- final_df_normalize3[train.index,]
valid.df <- final_df_normalize3[-train.index,]
```




```{r}

reg_kmeans_H <- lm(avg_meter_reading ~. - kmeans +factor(kmeans)     
                   -HierCluster + factor(HierCluster) 
                   + age*floor_count -avg_cloud - avg_sea -avg_wind, 
                   data = train.df)

summary(reg_kmeans_H)
vif(reg_kmeans_H)
```

```{r}
outlierTest(reg_kmeans_H)

plot(reg_kmeans_H)

reg_kmeans_H <- lm(avg_meter_reading ~. - kmeans      
                   -HierCluster + factor(HierCluster) 
                   + age*floor_count -avg_cloud - avg_sea -avg_wind, 
                   data = train.df[-c(13783,13683),])

summary(reg_kmeans_H)
vif(reg_kmeans_H)
```



```{r}
reg.train.predict <- predict(reg_kmeans_H, train.df, type = "response")
train.residuals <- train.df$avg_meter_reading - reg.train.predict
RMSEtrain <- sqrt(mean(train.residuals^2,na.rm=TRUE))

reg.valid.predict <- predict(reg_kmeans_H, valid.df, type = "response")
valid.residuals <- valid.df$avg_meter_reading - reg.valid.predict
RMSEvalid <- sqrt(mean(valid.residuals^2,na.rm=TRUE))

hist(valid.residuals, breaks = 25, xlab = "Residuals", main = "")

data.frame(actual = valid.df$avg_meter_reading, predicted = reg.valid.predict)


```
